# Handling Migration Data in R

## Contingency Table

### Contingency Table

- Bilateral migration flow data are commonly represented in square tables. 
- Values in non-diagonal cells represent a origin-destination count of migration between a specified set of regions. 
- Values in diagonal cells represent some form of non-moving population, or those that move within a region, which are typically not presented. 

<table>
    <tr>
        <td colspan=1>*Origin*</td>
        <td colspan=5 style="text-align:center">*Destination*</td>
    </tr>
    <tr style="border-bottom:2px solid black">
        <td></td>
        <td>A</td>
        <td>B</td>
        <td>C</td>
        <td>D</td>
        <td>Sum</td>
    </tr>
    <tr>
        <td>A</td>
        <td></td>
        <td>100</td>
        <td>30</td>
        <td>70</td>
        <td>200</td>
    </tr>
    <tr>
        <td>B</td>
        <td>50</td>
        <td></td>
        <td>45</td>
        <td>5</td>
        <td>100</td>
    </tr>
    <tr>
        <td>C</td>
        <td>60</td>
        <td>35</td>
        <td></td>
        <td>40</td>
        <td>135</td>
    </tr>
    <tr>
        <td>D</td>
        <td>20</td>
        <td>25</td>
        <td>20</td>
        <td></td>
        <td>65</td>
    </tr>
    <tr>
        <td>Sum</td>
        <td>130</td>
        <td>160</td>
        <td>95</td>
        <td>115</td>
        <td>500</td>
    </tr>
</table>


<!-- ```{=latex} -->
<!-- \begin{tabular}{lrrrrr}																											 -->
<!-- \toprule																											 -->
<!-- \cmidrule{1-6}																											 -->
<!-- \textit{Origin}  	&	 \multicolumn{5}{c}{\textit{Destination}} 									\\																 -->
<!-- \cmidrule{2-6}																											 -->
<!-- \textit{ }  	&	 A 	&	 B 	&	 C 	&	 D 	&	 Sum 	\\																 -->
<!-- \cmidrule{1-6}																											 -->
<!-- A 	&		&	100	&	30	&	70	&	200	\\																 -->
<!-- B 	&	50	&		&	45	&	5	&	100	\\																 -->
<!-- C 	&	60	&	35	&		&	40	&	135	\\																 -->
<!-- D 	&	20	&	25	&	20	&		&	65	\\																 -->
<!-- Sum	&	130	&	160	&	95	&	115	&	500	\\																 -->
<!-- \bottomrule																											 -->
<!-- \end{tabular}																											 -->
<!-- ``` -->

### Contingency Table

- Often denoted as $m_{ij}$
    - Row totals, the out-migration counts: $\sum_j m_{ij} = m_{i+}$
    - Column totals, the in-migration counts: $\sum_i m_{ij} = m_{+j}$
    - Net migration totals: $m_{i+} - m_{i+}$
    - Total migration: $m_{++}$
    
## Data Manipulation


### R `matrix` and `array`


- Some functions for describing and estimating migration in R require flow tables as `matrix` or `array` type objects
- Create a `matrix` in R using the `matrix()` function
    - Data read in by column. Change change using `byrow = FALSE`
    - Use the `dimnames` argument to supply region names

```{r}
# create region labels
r <- LETTERS[1:4]
r

# create matrix
m0 <- matrix(data = c(0, 100, 30, 70, 50, 0, 45, 5, 60, 35, 0, 40, 20, 25, 20, 0), 
             nrow = 4, ncol = 4, byrow = TRUE,
             dimnames = list(orig = r, dest = r))
m0
```

### R `matrix` and `array`

- Create an `array` in R using the `array()` function

```{r}
m1 <- array(data = sample(x = 1:100, size = 32), 
            dim = c(4, 4, 2), 
            dimnames = list(orig = r, dest = r, sex = c("female", "male")))
m1
```

### Show totals

- The `addmargins()` functions adds extra row, column and tables to display the dimension sums.
```{r}
addmargins(A = m0)
```



### Convert to matrix

- Data will not always come as an `matrix` or an `array`.
- There a couple of useful functions in R to convert data to when working with migration tables in R
- The `xtab()` function converts data frames into a `matrix` or `array`
    - `formula` column names with 
        - left hand side the column name to fill the `matrix` or `array`
        - a `~` to separate the left and right hand side
        - right hand side the columns to cross-classifying the left hand variable (separated by `+`).
    - `data` containing the variables for `formula`
- The `as.data.frame.table()` function takes a `matrix` or `array`  and converts it to a `data.frame` based on the array dimension names.
    - `responseName` to set the column name of based on the cells of the `matrix` or `array`


### Convert to matrix

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
d0 <- expand_grid(orig = r, dest = r) %>%
  mutate(flow = 1:16)
```

```{r}
# tidy migration data
d0
```

### Convert to matrix

```{r}
# convert to matrix
m2 <- xtabs(formula = flow ~ orig + dest, data = d0)
m2
```


### Convert to data frame

```{r}
# convert back to tibble
m2 %>%
  as.data.frame.table(responseName = "flow") %>%
  as_tibble()
```

### Convert to data frame

```{r}
# convert array to tibble
d1 <- m1 %>%
  as.data.frame.table(responseName = "flow") %>%
  as_tibble()
d1
```

## Matrix Operations

### Displaying migration matrics

- When dealing with migration `matrix` objects in R, they often are difficult to view
    - Lengthy dimension names, 
    - Unit size
    - Diagonal terms included but not of interest
- Some helpful R functions to adapt objects for easier viewing
- Demonstrate with the `uar_1960` object in the *migest* package      
    - Lifetime migration matrix for Governorates of United Arab Republic in 1960 used in the manual of @UnitedNations1983

### Displaying migration matrics

```{r}
library(migest)
uar_1960
```

### Abbriviate names

- View and alter the `matrix` dimension names using `rownames()` and `colnames()` or `dimnames()`
- The `abbreviate()` function applies an algorithm to shorten names
```{r}
dimnames(uar_1960)
# make a copy
u0 <- uar_1960
# new abbreviated region names
r <- list(orig = uar_1960 %>%
            rownames() %>%
            abbreviate(),
          dest = uar_1960 %>%
            colnames() %>%
            abbreviate())
```

### Abbriviate names

```{r}
r
# apply the abbreviated region names
dimnames(u0) <- r
```


### Abbriviate names

```{r}
u0
```

### Data scaling

- Basic arithmetic operators to scale the data to an appropriate level
- The `round()` function to specify precision of numbers
```{r}
u1 <- round(x = u0/1000, digits = 1)
u1
```

### Diagonal elements

- Set diagonal terms (non-movers) to zero using the `diag()` function
```{r}
u2 <- u0
diag(u2) <- 0
u2
```


## Summaries

### Net flows and counterflows

- The *migest* package contains a number of functions to provide summaries of origin-destination migration data
- The `counter()` function calculates the counter flow and net flow
    - Accepts `matrix` or `data.frame` (or `tibble`) inputs

```{r}
counter(m0)
```
### Net flows and counterflows

```{r}
d1 %>%
  group_by(sex) %>%
  counter()
```


### Totals

- The `sum_turnover()` provides summary in-migration, out-migration, net-migration and turnover totals for each region
    - Accepts `matrix` or `data.frame` (or `tibble`) inputs
    - Setting `type = "international"` to change labels in outputs

```{r}
sum_turnover(m0)
```

### Totals

```{r}
sum_turnover(m = d0, type = "international")
```


### Totals

- The `sum_turnover()` function can be applied with to large data sets spanning multiple years (groups)
- Demonstrate using international flow estimates of @Abel2019
```{r, cache = TRUE, message=FALSE, warning=FALSE}
# read data from web depository
f <- read_csv("https://ndownloader.figshare.com/files/26239945")
f
```


### Totals

```{r, cache = TRUE}
# single period
f %>% 
  filter(year0 == 1990) %>%
  sum_turnover(flow_col = "da_pb_closed", type = "international")
```

### Totals

```{r, cache = TRUE}
# all periods using group_by
f %>% 
  group_by(year0) %>%
  sum_turnover(flow_col = "da_pb_closed", type = "international") %>%
  arrange(country)
```

### Rest of categroies

- The `sum_lump()` function can be used to aggregate up smaller regions.
    - Specify the the desired level of small flows using the `threshold` argument 
    - Specify the `lump` argument to apply the `threshold` argument to either the `flow` values or the `in` and `out` totals.

```{r}
m0

# threshold on flows (default)
sum_lump(m0, threshold = 50)
```

### Rest of categroies


```{r}
addmargins(m0)
# threshold on in and out totals
sum_lump(m0, threshold = 120, lump = c("in", "out"))
```

### Rest of categroies

- Useful to reduce the number of corridors when plotting large data sets:

```{r, warning=FALSE}
# add continental regions to the global flow data set
library(countrycode)
d <- f %>% 
  filter(year0 == 2015) %>%
  mutate(
    orig_reg = 
      countrycode(sourcevar = orig, origin = "iso3c", dest = "un.region.name"),
    dest_reg = 
      countrycode(sourcevar = dest, origin = "iso3c", dest = "un.region.name")) %>%
  relocate(contains("orig"), contains("dest"))
d
```

### Rest of categroies

- Apply the `sum_lump()` function to lump together smaller flows (less than 100,000) within and between continents.

```{r, warning=FALSE}
d %>%
  group_by(orig_reg, dest_reg) %>%
  sum_lump(threshold = 1e5, flow_col = "da_pb_closed")
```



### Exercise (ex2.R)


```{r eval = FALSE, prompt=FALSE,  code = readLines('exercise/ex2.R') }
```
