# Summary Migration Indices

## Background

### Background

- Compared with fertility and mortality, little attention given to the way that internal or domestic migration varies between nations.
- Comparisons of migration over time or between spatial units is complicated by many factors including:
    - Different or changing definitions of migration
    - Different or changing collection systems for migration
    - Different sizes of regions 
    - Different or changing number of regions
    - Different ways of measuring distances
    - Different and changing underlying population sizes and structures
- @Bell2002 brought together and proposed a number of summary measures to enable better comparisons
    - A number of these incorporated into the IMAGE software for comparisons of internal migration between many countries; @Bell2015, @Bell2015a, @Stillwell2016, @Rees2016, @Bernard2014
    

### Background

- @Bell2002 identified four main groups of migration indices:
    - Intensity of migration
    - Distance of migration
    - Migration connectivity
    - Effect of migration on the redistribution of populations

## Intensity

### Migration intensity

- Migration intensity measures attempt to capture the overall level, or incidence, of mobility.
- Provide a single measure for comparison of migration intensities over time or space
- Some indices based on age-specific migration data
    - Will discuss later in the age-schedule section
    
### Crude migration probability

- Crude migration intensity is a simple measure of the overall propensity to migrate
    - Similar to crude birth or death rate
- If using migration transition data, the crude migration probability (CMP) is
$$
\texttt{CMP} = 100 M/P
$$
where $M$ is the total number of migrants in a given time period and $P$ is the population at risk

### Migration intensity

- @Courgeau1973 discussed how the intensity of migration is directly related to the number of regions $n$ in the country
$$
\texttt{CMP} = k \log (n^2)
$$
- No intrinsic meaning to a single Courgeau's $k$ , but can be used to compare migration intensity that cannot be seen from the raw data because of differences in their zonal systems.
- Higher the value of $k$, greater the intensity of migration

### Migration intensity

- The `index_intensity()` function in the *migest* package calculates both intensity measures, given a migration and population data 
- The *migest* package also contains a data set on Korean internal migration and populations of first level administrative districts
    - Data originally downloaded from `https://kosis.kr/eng`

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(migest)
korea_reg
```
### Migration intensity

- The `korea_pop` contains the resident population in each region and year
```{r}
korea_pop
```

### Migration intensity

- Calculate migration and population totals in 2020
```{r, warning=FALSE, message=FALSE}
m <- korea_reg %>%
  filter(year == 2020,
         orig != dest) %>%
  pull(flow) %>%
  sum()
m
p <- korea_pop %>%
  filter(year == 2020) %>%
  pull(population) %>%
  sum()
p

index_intensity(mig_total = m, pop_total = p,
                n = n_distinct(korea_pop$region))
```


### Migration intensity

```{r, warning=FALSE, message=FALSE}
mm <- korea_reg %>%
  group_by(year) %>%
  filter(orig != dest) %>%
  summarise(m = sum(flow))
mm
```

### Migration intensity

```{r, warning=FALSE, message=FALSE}
pp <- korea_pop %>%
  group_by(year) %>%
  summarise(p = sum(population))
pp
```

### Migration intensity

- Passing the vectors of migration and population totals can lead to confusing output
```{r, warning=FALSE, message=FALSE}
d <- mm %>%
  left_join(pp)
d
```

### Migration intensity

```{r, warning=FALSE, message=FALSE}
index_intensity(mig_total = d$m, pop_total = d$p,
                n = n_distinct(korea_pop$region))
```

### Migration intensity

- Set `long = FALSE` to put each indicator in their own column
```{r, warning=FALSE, message=FALSE}
index_intensity(mig_total = d$m, pop_total = d$p,
                n = n_distinct(korea_pop$region), long = FALSE)
```

### Migration intensity

- Use the `map2()` function in *purrr* to apply the function to subsets of the data
    - Provides results alongside the `year`
- Code below produces a nested list `i` containing the intensity measures for each year
```{r, warning=FALSE, message=FALSE}
d <- mm %>%
  left_join(pp) %>%
  mutate(i = map2(.x = m, .y = p,
                  .f = ~index_intensity(mig_total = .x,
                                        pop_total = .y,
                                        n = n_distinct(korea_pop$region),
                                        long = FALSE))) 
d
```

### Migration intensity

- The `unnest()` function in the *tidyr* package binds each component of column `i` on top of each other
    - Easier to see changes over time
```{r}
unnest(d, cols = i)
```



## Distance

### Migration distance

- As migration is a spatial activity, based on movements between two locations, comparisons should take account of the way that intensities of movement vary across space.
- There are a number of measures that summarize the effects of distance across a migration system
- The distance measure between each region is not straightforward
    - Ideally measure the typical distance that migrants travel.
    - Straight line distance between population-weighted centroids of each region provide a good approximation
- The costs faced by a migrant may not be represented well by the inter-centroid distance. 
    - Locations in different regions might be very close to a border, so centroids will exaggerate the distance 
    - Areas can take many shapes and sizes
    - Doughnut shaped regions might have centroids not in region
    - Indented coastlines might make regions look closer than they might be (culturally, travel cost)
    - Distance measures for within region moves cause another set of problems

### Average migration distance

- Summary of the average migration distance can be calculated by taking a weighted average of the migration counts, where the corresponding distances are the weights
- @Bell2002 note a median average as clearly preferable to a mean average as the distribution of distances is negatively skewed, reflecting the strong distance decay effect which consistently occurs
- Comparison with the mean average distance provides a guide to the skewness

### Distance decay

- A more complete method to account for the skewness in migration distances is to fit a model to predict migration counts using the distances between each regions and extract the distance parameter
- Different models could be potentially used, but tend to be based on a `log(distance)` terms with categorical control variables for the origin and destination
- The distance decay parameter ($\beta_3$) in a Poisson log-linear model; 
$$
\log(m_{ij}) = \beta_0 + \beta_{1i} \texttt{origin} + \beta_{2j} \texttt{destination} + \beta_{3} \log(\texttt{distance})
$$
- The distance decay parameter of interest ($\beta_3$ in the equation above) is almost always negative, indicating an increase in migration leads to fewer predicted migrations.
    - The set of $\beta_{1i}$ and $\beta_{2j}$ represent some form of push and pull factors for each region ($i$ and $j$)

### Migration distance

- The `korea_dist` matrix provides estimates of the 2020 population weighted distances in kilometers between the 17 first level administrative districts in Korea
    - Data based on WorldPop estimates of @Edwards2021
```{r}
korea_dist
```

### Migration distance

- The `index_distance()` function in the *migest* package provides three summary distance measures given a set of migration and distance measures between each origin and destination. 
- The origin-destination migration flows can be given as a matrix to `m` or as a data frame, where the column names are assumed to be `orig`, `dest` and `flow`. 
    - Can change using `orig_col`, `dest_col` and `flow_col` arguments
- The distance values can also be given as a matrix to `d` or as a data frame, where the column names are assumed to be `orig`, `dest` and `dist`. 
- Origin and destination names in `m` and `dist` must match
- Removes all within migration moves from calculations
    
```{r}
# single year
index_distance(m = filter(korea_reg, year == 2020),
               d = korea_dist)
```


### Migration distance

```{r}
korea_reg %>%
  nest(m = c(orig, dest, flow)) %>%
  mutate(d = list(korea_dist)) %>%
  mutate(i = map2(.x = m, .y = d,
                  .f = ~index_distance(m = .x, d = .y, long = FALSE))) %>%
  unnest(i)
```



### Calculating distances

- There are a number of functions to a calculate distance matrices in R
- Require a set of longitude and latitudes 
- If population weighted centriods are not available from national statistics offices, a number of research centers provide estimates
    - POPGRID Data Collaborative `https://www.popgrid.org/`
- Example for Ghana using WorldPop 2020 population weighted centroids
    - CSV from `https://www.worldpop.org/doi/10.5258/SOTON/WP00703`

```{r, message=FALSE, warning=FALSE}

g <- read_csv("data/PWD_2020_sub_national_100m.csv") %>%
  filter(ISO == "GHA")
g
```

### Calculating distances


```{r}
g <- g %>%
  filter(ISO == "GHA") %>%
  select(Adm_N, PWC_Lon, PWC_Lat)
g
```


### Calculating distances

- The `distm()` function in the *geosphere* package provides great circle distance estimates in meters between centroids
```{r}
library(geosphere)
ghana_dist <- g %>%
  select(PWC_Lon, PWC_Lat) %>%
  distm()

ghana_dist
```


### Calculating distances

- Add the origin and destination names to the `matrix` row and columns using `dimnames()`
- Allows to combine with migration data
- Divide by 1000 to get to kilometers
```{r}
dimnames(ghana_dist) <- list(orig = g$Adm_N, dest = g$Adm_N)
round(ghana_dist/1000)
```


## Connectivity

### Migration connectivity

- The size of migration flows in different migration corridors vary due to many reasons other than population sizes and distance.
    - Also reflect the strength of many other factors linking regions such as the strength of historical, cultural, social and economic ties, between regions. 
- @Bell2002 note a fragmentation in the literature on measures of connections in a migration system and the use of a range of terms including *spatial connectivity*, *spatial concentration*, *spatial inequality* and *spatial focusing*
- There are many indices on migration connectivity. See the *migration.indices* package for example.
- The `index_connectivity()` function provides 12 different measures, that can broadly be placed into 5 groups.
    - Requires a matrix or data frame of migration flows
    - When providing a data frame, function is assuming flows are in a column with name `flow`; change with `flow_col` if not.
    
### Migration connectivity

```{r}
korea_reg %>%
  filter(year == 2020) %>%
  index_connectivity()
```
### Connectivity and Inequality

- The `connectivity` measure evaluates the proportion of the flows (excluding within region flows) that are non-zeros
    - More useful when many regions where populations are smaller
- @Bell2002 inequality measures are based on a distributions of flows compared to distributions of expected flows
    - `inequality_equal` measures the distance of the observed flows to an expected distribution where all flows are equal
    - `inequality_sim` measures the distance of the observed flows to an expected distribution from a spatial interaction model equivalent to a Poisson regression mode for an independence fit
$$
\widehat{\log(m_{ij})} = \beta_0 + \beta_{1i} \texttt{origin} + \beta_{2j} \texttt{destination}
$$
- In both cases, a value of 0 shows the observed flows match the expected values (some form of equality) and 1 shows the maximum distance between the observed flows and the expected flows, i.e. maximum inequality.

### Gini measures

- The Gini measures provide a value for the spatial focusing in a set of migration flow - i.e. how much of the migration is focused on a particular set of migration corridors
- Compares each flow with every other flow in the migration matrix. 
- A `gini_total` value of zero indicates all flows are of equal size (no spatial focusing) to 1, only one single flow (maximum focusing).
- The `gini_orig_standardized` values provide a similar measure but compare every outflow from each origin with every other outflow from that origin.
    - Measures the extent to which the destination choices of out-migrants are spatially focused. 
    - The `gini_dest_standardized` does the same but for the spatial focusing of origins of in-migrants.
    - The standardized values ensure a range of 0 and 1 - zero is no focusing

### Migration totals Gini and Coefficient of Variation

- The migration weighted Gini indexes provide a measure of the focusing for the in-migration and out-migration totals (`mwg_orig` and `mwg_dest`)
- The `mwg_mean` is a simple average of `mwg_orig` and `mwg_dest` to provide a system wide measure of focusing for all migration totals.
    - As with the `gini_` measures from `index_connectivity()` values vary between zero (no focusing) and 1 (all migration goes through a single origin or destination)
- @Rogers1998 proposed a coefficient of variation, provided by `cv` which compares the mean of the flows to the standard deviation of the flows.
    - Is not limited to 0 and 1
- The `acv` provides a similar measures of variation but based on the aggregate of coefficient of variations of in- and out-migration totals (based on the means and standard deviations of in- and out-migration totals)
    - Again, not limited to 0 and 1, but useful for comparisons across time or countries, where rising `cv` or `acv` would indicate greater inequality in migration flows or flow totals


## Impact

### Migration impact

- The impact of migration measures the extent to which migration acts to transform the pattern of human settlement
- Migration is already or becoming predominant mechanism leading to the redistribution of population in many regions of the world
- Descriptive studies tend to focus on regional net migration patterns
- Additional measures exist that summarize the overall effect of migration in redistributing a population across the entire system of regions

### Migration effectiveness index

- The migration effectiveness index (MEI) of @Shryock1976 compares the sum of net migration as a proportion of migration turnover, measuring the amount asymmetry or equilibrium in the migration network
$$
\texttt{MEI} = 100 \frac{\sum_{i} | \texttt{net} |}{\sum_{i} \texttt{turnover}_i} = 100 \frac{\sum_{i} | m_{+i} - m_{i+}|}{\sum_{i} m_{+i} + m_{i+}}
$$
- MEI range between 0 and 100. 
- High values indicate migration is an efficient mechanism of population redistribution, based on large net totals for the given turnover.
- Values closer to zero are generated from more balanced migration systems with less population redistribution

### Aggregate net migration rate

- The aggregate net migration rate (ANMR) of @Bell2002 attempts to measure the overall effect of migration on the population settlement patterns by replacing the denominator of $MEI$ with the each regions population $P_i$
- Index measures the net shift of population between regions per 100 residents in the country
    - No upper limit
$$
\texttt{ANMR} = 100 \frac{1}{2} \frac{\sum_{i} |\texttt{net}_i|}{\sum_{i} P_i} = 100 \frac{1}{2} \frac{\sum_{i} | m_{+i} - m_{i+} |}{\sum_{i} P_i}
$$

- Product of the CMI and MEI


### Preference and velocity

- The manual by @UnitedNations1983 provides two other impact measures, that seem to have fallen out of favor
- The preference index is based on an expected model of migration intestines based on population shares and the overall level of migration $M \frac{p_{i}}{P}\frac{p_{j}}{P}$, where $M$ is the total migration flow and $P$ is the total population based on the sum of populations in each region ($p_{i}$)
- Index compares the observed flows to an expected model where migration rates in all populations are the same
    - No upper limit
$$
\texttt{preference} = \sum_{ij}{\frac{m_{ij}}{M \frac{p_{i}}{P}\frac{p_{j}}{P}}}
$$

- The velocity index is based on a migration velocity measure $\frac{m_{ij}}{p_{i} p_{j}}$, multiplied by the total population and summed
- Index compares observed flows to an expected models where flows sizes are determined by population sizes alone
    - No upper limit
$$
\texttt{velocity}  = \sum_{ij}{\frac{m_{ij}}{p_{i} p_{j}}{P}}
$$

### Migration impact

- The `index_impact()` function in *migest* calculates all four measures given a set of migration flows and population sizes in each region
    - The `p` parameters assumes column names `region` and `pop` for region and population. Change from defaults using `reg_col` and `pop_col`
```{r}
index_impact(
  m = subset(korea_reg, year == 2020),
  p = subset(korea_pop, year == 2020),
  pop_col = "population"
)
```


### Migration impact

- Multiple years require nesting the migration and population data bases besides each other
```{r}
d <- korea_reg %>%
  nest(m = c(orig, dest, flow)) %>%
  left_join(korea_pop) %>%
  nest(p = c(region, population))
d
```

### Migration impact

- Apply the `index_impact()` function to each row and unnest
```{r}
d %>%
  mutate(i = map2(.x = m, .y = p, 
                  .f = ~index_impact(m = .x, p = .y, 
                                     pop_col = "population",
                                     long = FALSE))) %>%
  unnest(i)
```

### Exercise (ex3.R)

```{r eval = FALSE, prompt=FALSE,  code = readLines('exercise/ex3.R') }
```
