---
title: Describing Bilteral Migration Data
author: Guy J. Abel
date: 
fontsize: 10pt
output:
  beamer_presentation:
    theme: "Frankfurt"
    slide_level: 3
    keep_tex: true
    includes:
      in_header: header.txt
bibliography: C:\\Users\\Guy\\Documents\\library.bib      
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(prompt= TRUE, collapse = TRUE, comment = NA)
knitr::knit_hooks$set(purl = hook_purl)

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Components

## .

### Multiplicative Component Model 

-  @rogers2002dms proposed dis-aggregating origin-destination flow tables into separate components to allow for an easier examination of migration flows
    - Overall component - level of migration $\gamma$
    - Origin component - relative ‘pushes’ from each region $\alpha_i$
    - Destination component - relative ‘pulls’ to each region $\beta_j$
    - Origin–Destination interaction component - physical or social distance between places not explained by the overall and main effects. $\delta_{ij}$
- Simple calculations to estimate each component: 
$$
\gamma = m_{++} \qquad \alpha_i = \dfrac{m_{i+}}{m_{++}} \qquad \beta_j = \dfrac{m_{+j}}{m_{++}} \qquad  \delta_{ij} = \dfrac{m_{ij}}{\gamma\alpha_i\beta_j}
$$
- The interaction, $\delta_{ij}$, is the ratio of observed flow to an expected flow (for the case of no interaction). 

### Multiplicative Component Model 


- The dis-aggregation of the components is multiplicative: 
$$
m_{ij} = \gamma \alpha_i \beta_j \delta_{ij}
$$
- Equivalent to a saturated Poisson regression model ($R^2 = 1$)  where 
    - $\gamma$ is constant term
    - $\alpha_i$ is categorical term for the origin regions
    - $\beta_j$ is categorical term for the destination regions
    - $\delta_{ij}$ is an interaction term between the  $\alpha_i$ and  $\beta_j$
$$
\log m_{ij} = \gamma +  \alpha_i ORIG_i + \beta_j DEST_j +  \delta_i ORIG_i:DEST_j
$$
- When data is in a tidy format with row $h$ would be:
$$
\log y_{h} =  \beta_0 +  \beta_1 ORIG_{h} + \beta_2 DEST_{h} + \beta_3 ORIG_{h}:DEST_{h}
$$
- Poisson regression models such as these - where all the predictor variables are categorical - are also know as log-linear models
- Standard functions for fitting regression models, such as `glm()` in R will provide the same fitted values, but different parameter estimates
    - Use different coding system for the constraints when estimating parameters
    - Rogers' terms the parameter estimates using the equations for $\gamma, \alpha_i, \beta_j$ and $\delta_{ij}$ above the *total reference* coding system

### Multiplicative Component Model 

- The *migest* package contains a `multi_comp()` function to generate parameter estimates from an origin-destination flow matrix
    - Demonstrate with previous dummy data set

```{r, message=FALSE}
r <- LETTERS[1:4]
m0 <- matrix(data = c(0, 100, 30, 70, 
                      50, 0, 45, 5, 
                      60, 35, 0, 40, 
                      20, 25, 20, 0), 
             nrow = 4, ncol = 4, byrow = TRUE, 
             dimnames = list(orig = r, dest = r))
addmargins(m0)
```

### Multiplicative Component Model 

```{r, message=FALSE, message=FALSE}
library(tidyverse)
library(migest)
m0 %>%
  multi_comp() %>%
  round(3)
```

### Multiplicative Component Model 

- As the model is saturated, the fitted values are the same as the observed values.

```{r, message=FALSE}
multi_comp(m = m0)

# fitted value for A to B
500 * 0.4 * 0.32 * 1.5625
```


### Multiplicative Component Model 

- The total reference coding scheme for the parameter estimates are easier to examine than parameter estimates from a Poisson model fitted using `glm()`
    - More detail on `glm()` in next section
```{r}
d0 <- as.data.frame.table(x = m0, responseName = "flow")
f0 <- glm(formula = flow ~ orig + dest + orig * dest, family = poisson(), 
           data = d0)
f0
```

### Multiplicative Component Model 

```{r}
# fitted and observed values are the same
d0 %>% 
  as_tibble() %>%
  mutate(fit = round(f0$fitted.values, digits = 5))
```


### Multiplicative Component Model 

- Rogers' and colleagues have used the multiplicative component model to estimate migration flow tables
- Expand to multiple dimensions
- Rectify bumpy age schedules
    - Replace reported age parameters (proportions) in the multiplicative component model with proportions from a more regular schedule. 
    - Multiply the new age parameters with the existing total, origin, destination and interaction parameters to obtain new estimated flows. 

### Multiplicative Component Model

- Italian data in *migest* package
```{r, message = FALSE}
italy_area
```

### Multiplicative Component Model

```{r}
# single year, multiple age groups
c0 <- italy_area %>%
  filter(year == 2000) %>%
  multi_comp()
round(c0, 3)
```

### Multiplicative Component Model

```{r}
# origin components (shares)
c0 %>%
  as.data.frame.table(responseName = "comp") %>%
  filter(orig != "Sum", dest == "Sum", age_grp == "Sum")

# destination components (shares)
c0 %>%
  as.data.frame.table(responseName = "comp") %>%
  filter(orig == "Sum", dest != "Sum", age_grp == "Sum")
```

### Multiplicative Component Model

```{r, fig.height=4}
# age components
c0 %>%
  as.data.frame.table(responseName = "comp") %>%
  filter(orig == "Sum", dest == "Sum", age_grp != "Sum") %>%
  ggplot(mapping = aes(x = age_grp, y = comp, group = 1)) +
  geom_line() +
  theme(axis.text = element_text(angle = 45, hjust = 1))
```

# Log-Linear Models

## .

### Log-Linear Models

- Rogers' and collaborators like to shorten the multiplicative form of the log-linear model to use capital letters to represent parameters
$$
m_{ij} = \gamma \alpha_i \beta_j \delta_{ij} = T O_i D_j OD_{ij}
$$
- When there is multiple origin-destination tables, by different age groups, sex, education level, etc,... the notation can be easily used to study different log-linear models 
$$
m_{ij} = T O_i D_j A_x OD_{ij} OA_{ix}
$$
- When data is in a tidy format with row $h$ would be:
$$
\begin{aligned}
\log y_{h} =&  \beta_0 +  \beta_1 ORIG_{h} + \beta_2 DEST_{h} + \beta_3 AGE_{x} + \\
&\beta_4 ORIG_{h}:DEST_{h} + \beta_5 ORIG_{h}:AGE_{h} 
\end{aligned}
$$


### Log-Linear Models


- We can fit log-linear models in R using the `glm()` function (for generalised linear models)
- Requires a `formula`, `data` and `family` argument
- The `formula` argument is similar to that in `xtabs()`, where we use the `~` symbol to separate the the response and explanatory variables
    - For example the model in the previous slide would use `formula = flow ~ orig + dest + age + orig:dest + orig:age`
    - Use `:` or `*` to denote interactions
- The `family` argument should be set to `poisson()` for a log-linear model

### Log-Linear Models

- Example with age-specific migration flows between Italian regions in 1970

```{r}
d1 <- italy_area %>%
  filter(orig != dest, 
         year == 1970) %>%
  # rename so later output fits on slide
  rename(age = age_grp)
d1
```

    
### Log-Linear Models

```{r}
glm(formula = flow ~ orig + dest, family = poisson(), data = d1)
```

# Dimensions

## . 

### Log-Linear Model Analysis

- As we increase the number of dimensions of the data, it might become important to understand which dimensions of the data are most important
- We can use log-linear models with detailled migration data to
    - Understand the dominate dimensions, for example @VanImhoff1997 @rogers2002dms
    - Predict origin-destination flows with partial data, for example @deBeer2010 @Rogers2003a @Raymer2007b
    - Project detailed origin-destination flows, for example @Raymer2006
- All the above examples involve fitting a number log-linear models based on different dimensions of the data frames
    - Use model fit statistics to judge the best model
    
### Log-Linear Model Analysis

- One approach to choosing the most important dimensions is to fit all possible combinations of models - known as *dredging* the model space
- The `dredge()` function in the *MuMIn* package will fit all combinations of regression models given an upper limit, i.e. the most complex model.
    - The number of combinations grows exponentially with the number of predictors 
    - Does not allow `na.action = "na.omit"` - set by default in `glm()` for handling missing values in regression models


### Log-Linear Model Analysis

- Fit the most complex model using `glm()`. 
    - Set `na.action = na.fail` to exclude failed models in when using the `dredge()` function later
    - Most complex model typically involves at least all two-way interactions
- The `formula` argument in `glm()` allows the use `()^2` to construct all two-way interactions, i.e. the below give the identical outputs
    - Use `()^3` for all three way interactions
```{r, cache=TRUE}
f1 <- glm(formula = flow ~ (orig + dest + age)^2, 
          family = poisson(), data = d1, na.action = na.fail)
f2 <- glm(formula = flow ~ orig * dest + orig * age + dest * age,
          family = poisson(), data = d1, na.action = na.fail)

# check terms used in models
attr(f1$terms, "term.labels")
attr(f2$terms, "term.labels")
```

### Log-Linear Model Analysis

- Models will have many estimated coefficients 
    - Some will be non-determinable because no observations (e.g. diagonal terms such as `origIslands:destIslands` below) as 
```{r, cache=TRUE}
f1 %>% 
  coef() %>% 
  length()
summary(f1)
```



### Log-Linear Model Analysis

- Pass the upper model to `dredge()`. Use `trace = TRUE` to monitor progress.
```{r, cache=TRUE}
library(MuMIn)
mm <- dredge(global.model = f1, trace = TRUE)
```

### Log-Linear Model Analysis


```{r}
mm
```


### Log-Linear Model Analysis

- Model comparison based on model statistics measuring the goodness of fit.
    - AIC measures a goodness of fit with a penalty for the number of predictor variables. 
    - AICc has a bias correction term for small samples
- Typically the origin-destination interaction term is very important for accurately predicting the age-specific origin-destination migration flows
- The time to conduct a dredging analysis increase exponentially as the number of dimensions increases.

### Exercise (ex6.R)

```{r eval = FALSE, prompt=FALSE,  code = readLines('../exercise/ex6.R') }
```




### References  {.allowframebreaks} 

\scriptsize