---
title: Chord Diagrams for Visualising Bilateral Migration 
author: Guy J. Abel
date: 
fontsize: 10pt
output:
  beamer_presentation:
    theme: "Frankfurt"
    slide_level: 3
    keep_tex: true
    includes:
      in_header: header.txt
bibliography: C:\\Users\\Guy\\Documents\\library.bib
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(prompt= TRUE, collapse = TRUE, comment = NA, cache = FALSE)
knitr::knit_hooks$set(purl = hook_purl)
```

# Background

## .

###  Background

- Visualizing bilateral migration is not straightforward
    - Difficult to represent the geographic and temporal aspect at the same time
- Many different approaches
    - Difficult to satisfy everyone's tastes
- In this class will illustrate two non-map based approaches
    - Chord Diagrams
    - Alluvial or Sankey Plots
- Non-map based approaches
    - Provide clearer visual guide for geographically small areas that can be overwhelmed on a map
    - Include more bilateral connections

###  Map Based - The Emigrants of the World, Minard 1858.

\begin{center}
  \includegraphics[width=0.8\textwidth]{./img/minard.png}
\end{center}

###  Flowline Maps

\begin{center}
  \includegraphics[width=\textwidth]{./img/lambert.png}
\end{center}


###  Criticised New York Times refugee flow map

\begin{center}
  \includegraphics[width=0.8\textwidth]{./img/refugee_flow_map-africa.jpg}
\end{center}


### Martin Grandjean's attempt to rectify

\begin{center}
  \includegraphics[width=\textwidth]{./img/Refugees-world-930x523.png}
\end{center}


# Chord Diagram

## .

###  Chord Diagrams

- First chord diagrams introduced by Martin Krzywinski in 2007.
    - `https://www.nytimes.com/imagepages/2007/01/22/science/20070123_SCI_ILLO.html`
- Used to facilitate the identification and analysis of similarities and differences arising from comparisons of genomes
- Displays relationships between pairs of positions by the use of ribbons, which encode the position, size, and orientation of related genomic elements
- Developed into Circos software in Perl by @Krzywinski2009
    - `http://circos.ca/`

### New York Times 2007

\begin{center}
  \includegraphics[width=0.7\textwidth]{./img/nyt.jpg}
\end{center}

###  Chord Diagrams with Migration Data

- Interactive chord diagram plots introduced into rr.js library by Bostock
- First used to illustrate migration patterns by data journalist Chris Walker in 2013
    - Mapping America’s Restless Interstate Migration Without a Map `https://www.wired.com/2013/11/mapping-migration-without-a-map/`
- Does not show the direction of move until mouse-over.
- Nikola Sander adapted Circos software to add directional indicators for flows
    - First used in @Abel2014. *Quantifying Global International Migration Flows*. Science, 343 (6178).
    - Interactive version at `http://download.gsb.bund.de/BIB/global_flow/`

###  Chord Diagrams with Migration Data

\begin{center}
  \includegraphics[width=0.7\textwidth]{./img/walker.jpg}
\end{center}

###  Chord Diagrams with Migration Data

\begin{center}
  \includegraphics[width=0.7\textwidth]{./img/sander.jpg}
\end{center}

# circlize

## .

### Chord Diagrams in R

- Some drawbacks to the Circos based plots
    - Inflows plotted first on each sector
    - Chords for smaller flows overlap larger flows
    - Hides smallest flows
    - Not easy to detect direction of flows
    - Addition of direction arrows usually require some further touch using a second piece of software, e.g. Photoshop or Illustrator
        - Problematic for replicability
- In recent years a number of R packages that implement similar plots as the Circos software have appeared on CRAN
- The *circlize* R package by @Gu2014 is perhaps the most complete and accessible for non-genomic data
    - Built on base R graphics package
- Includes a `chordDiagram()` function
    - Extensive documentation of the `chordDiagram()` function in [Chapters 13-15](https://jokergoo.github.io/circlize_book/book/the-chorddiagram-function.html) of the circlize book.


### UN international migrant stock data 2020

```{r, message=FALSE}
library(tidyverse)
un <- read_csv(file = "../data/un_desa_ims_tidy.csv")
un
```

### UN international migrant stock data 2020

- Use continent to continent flows in 2020
```{r}
# codes for contents
cc <- c(903, 935, 908, 904, 905, 909)
d <- un %>%
  filter(por_code %in% cc, 
         pob_code %in% cc,
         year == 2020)
d
```

### UN international migrant stock data 2020

- Remove within continent stocks (will dominate the plot) and focus on inter-continent migrants
```{r}
d <- d %>%
  rename(orig = pob_name,
         dest = por_name) %>%
  filter(orig != dest) %>%
  select(-contains("code"))
d

```


### Default `chordDiagram()`

- The `chordDiagram()` function can take either a `matrix` or `data.frame` object as first argument `x` for the data.
- I prefer the latter as they are much easier to create and manipulate (using *dplyr* and other *tidyverse* packages).
    - When using a `data.frame`, the first three columns should correspond to the origin, destination and size of connection.
    - Columns can take any name, but must be in that order.
    - Will also work with `tbl_df` (tibble)
- Many options in `chordDiagram()`, that by default are not ideal for displaying migration data

```{r cd-default,  fig.width = 4, fig.height = 4, message=FALSE,  warning= TRUE,  fig.show='hide', fig.path="./plot/"}
library(circlize)
# first three columns not origin, destination, connection (in that order)
chordDiagram(x = d)
```

### Default `chordDiagram()` - incorrect column order

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-default-1.pdf")
```

### Default `chordDiagram()`

- Move the `orig`, `dest` and `stock` columns to the left of the data frame using the `relocate()` function in the *dplyr* package
```{r cd-col-order, fig.width = 4, fig.height = 4, message=TRUE,  warning= TRUE,  fig.show='hide', fig.path="./plot/"}
d <- relocate(d, orig, dest, stock)
d
chordDiagram(x = d)
```

### Default `chordDiagram()` - includes 2 numeric columns

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col-order-1.pdf")
```


### Default `chordDiagram()`

- Avoid chord link ends with unequal widths at each base by using only one numeric column in `d`
```{r cd-column3, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
d <- select(d, orig, dest, stock)
d
chordDiagram(x = d)
```


### Default `chordDiagram()` - only 3 columns

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-column3-1.pdf")
```


# Sectors

## .

### Sector Axis

- Edit the bilateral counts to a sensible scale to ensure the axis labels are legible.

```{r cd-axis, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
d <- mutate(d, stock = stock/1e6)
d
chordDiagram(x = d)
```

### Sector Axis

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-axis-1.pdf")
```


### Sector ordering

- Sector ordering is alphabetical by default
- Can specify order using `order` argument and pass a vector
- Try to order so that neighboring regions are next each other

```{r}
r <- tibble(reg = union(d$orig, d$dest))
r
```

### Sector ordering

```{r cd-sector, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
r <- r %>%
  mutate(reg_order = c(4, 3, 6, 1, 5, 2)) %>%
  arrange(reg_order)
r

# order sectors
chordDiagram(x = d, order = r$reg)
```

### Sector ordering

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-sector-1.pdf")
```

### Orientation and gaps

- The `circos.par()` function controls the overall layout parameters of the graphic display
- Use `circos.par()` to set
    - `gap.degree` the degree of gaps between sectors are set - default `gap.degree = 1`
    - `start.degree` the degree from three o'clock where the first sector appears - default `start.degree = 0`
- Anything set via `circos.par()` will be fixed for all remaining pots
- Reset to default graphic parameters using `circos.clear()` or overwrite with new `circos.par()`

```{r cd-gap, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
# increase gaps
circos.par(gap.degree = 5)
chordDiagram(x = d, order = r$reg)

# rotate
circos.par(start.degree = 90)
chordDiagram(x = d, order = r$reg)
```

### Bigger gap - `gap.degree = 5`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-gap-1.pdf")
```

### Alternative orientation - `start.degree = 90`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-gap-2.pdf")
```



# Colour

## .

### Sector colours

- Colours are randomly generated (will change every time you plot)
- Can set to a choice using either:
    - `grid.col` corresponding to sectors (regions/countries/areas)
    - `transparency` set by default to 0.5

```{r cd-col1, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
r <- r %>%
  mutate(col1 = c("black", "gold", "orange", "blue", "purple", "red"))
r

chordDiagram(x = d, order = r$reg, grid.col = r$col1)
```


### Sector colour - set `grid.col`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col1-1.pdf")
```

### Sector colour

- Can use the *RColourBrewer* package to generate palettes (maximum of 9 colours)
    - Based on `https://colorbrewer2.org/`

```{r cd-col2, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE, warning= FALSE,  fig.show='hide', fig.path="./plot/"}
library(RColorBrewer)
r <- r %>%
  mutate(col2 = brewer.pal(n = 6, name = "Set1"),
         col3 = c("Red", rep("Grey", times = 5)))
r

chordDiagram(x = d, order = r$reg, grid.col = r$col2)
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25)
chordDiagram(x = d, order = r$reg, grid.col = r$col3)
```

### Sector colour - set `grid.col`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col2-1.pdf")
```

### Sector colour - `transparency = 0.25`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col2-2.pdf")
```

### Secotr colour - set `grid.col` to highlight

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col2-3.pdf")
```

### Chord colours

- Chord colours follow the origin sector. We can specify different colours using
    - `col` corresponding to links (bilateral migration data)
    - `link.visible` will hide particular chords
```{r}
d <- d %>%
  # highlight Asia to Europe flows
  mutate(link_col1 = ifelse(test = orig == "ASIA" & dest == "EUROPE",
                            yes = "black", no = "grey"),
         # show only flows out or into Asia
         show_link = orig == "ASIA" | dest == "ASIA")
d
```


### Chord colours

- Pass the chord specific settings to `chordDiagram()`
```{r cd-col3, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
chordDiagram(x = d, order = r$reg,
             grid.col = r$col2, col = d$link_col1)
chordDiagram(x = d, order = r$reg,
             grid.col = r$col2, link.visible = d$show_link)
```

### Chord colours - setting `col` to column in bilateral data

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col3-1.pdf")
```

### Chord colours - setting `link.visible`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-col3-2.pdf")
```

# Chords

## .

### Direction

- Distinguish direction of bilateral link using
    - Different heights at the start and end of the chord links
    - Arrows
    - Combination of both
- Set in `chordDiagram()` using
    - `directional = 1` (from link goes from first to second column)
    - `direction.type` arguments
```{r cd-direction1, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
# drop link_col column
d$link_col1 <- NULL

# as used by Sander, default of direction.type = "diffHeight"
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1)

# default arrows are too much
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"))

# getting there...
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow")
```



### Direction - default `direction.type = "diffHeight"`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction1-1.pdf")
```

### `direction.type = c("diffHeight", "arrows")`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction1-2.pdf")
```

### Direction - adding `link.arr.type = "big.arrow"`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction1-3.pdf")
```

### Direction

- Connect the base of the link to the sector through combination of
    - Adjusting the difference in height between the beginning and end of chords
    - Removing padding between the axis and the grid (the inner circle where the chords are)
- Set the `diffHeight` argument to a negative number so that the start of the chord is longer than then end.
   - Removes the destination sector bars (chart junk IMO).
```{r cd-direction2, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
# extreme height difference
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight  = -0.2)

# height difference looks good
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight  = -0.05)
```

### Direction `diffHeight = -0.2`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction2-1.pdf")
```

### Direction `diffHeight = -0.05`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction2-2.pdf")
```


### Direction

- Set in the `track.margin` option of `circos.par()` to remove the padding
    - Default of `track.margin = c(0.01, 0.01)` for chord diagrams - one percent between label names and the axis, and one percent between the axis and the grid (the chords)

```{r cd-direction3, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
# set second margin to zero
circos.par(track.margin = c(0.01, 0))
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05)

# set second margin to -0.01 to get seamless overlap
circos.par(track.margin = c(0.01, -0.01))
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05)


```


### (last plot repeated) - `track.margin = c(0.01, 0.01)`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction2-2.pdf")
```

### Direction - `track.margin = c(0.01, 0)`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction3-1.pdf")
```

### Direction - `track.margin = c(0.01, -0.01)`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction3-2.pdf")
```


### Chord ordering

- Number of options in `chordDiagram()` to control the chord link order
    - `link.sort` sort the order the links from largest to smaller as the enter and exit the plot, by default `FALSE`
    - `link.largest.ontop` sort the order of the plotting of the links so that the smallest are given less prominence. By default `FALSE`, so plots the links in the last sector last and they appear more predominant
```{r cd-order, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
# sort links on sectors
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05,
             link.sort = TRUE)

# sort link plotting order
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05,
             link.sort = TRUE, link.largest.ontop = TRUE)
```

### Chord ordering  (last plot repeated)

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-direction3-2.pdf")
```


### Chord ordering - `link.sort = TRUE`

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-order-1.pdf")
```

### Chord ordering - adding `link.largest.ontop = TRUE`


\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-order-2.pdf")
```

# Labels

## .

### Labels

- Multiple options for the orientation of labels, set via
    - `inside`, `outside`, `clockwise`, `reverse.clockwise`, `downward`, `bending.inside` and `bending.outside`
    - Cannot pass to `chordDiagram()` so we have to first use `annotationTrack` option to only plot the `grid` (the chords) and `axis` (default for `annotationTrack = c("name", "grid", "axis")`)
- To add the labels we use the `panel.fun` argument in `circos.track()`.
    - Works like a for loop, cycling through each sector of the track (the circle)
    - For each sector we use `circos.text()` to add labels at a specified `x` and `y` location
    - Can also set the `facing` orientation of the labels as well as other options such as text size (`cex`) and colour (`col`)

```{r cd-label1, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
# drop the name labels
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05,
             link.sort = TRUE, link.largest.ontop = TRUE,
             annotationTrack = c("grid", "axis"))
```

### Labels

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-label1-1.pdf")
```


### Labels

- No room for labels. We can create this using the `preAllocateTracks` argument.
    - Requires a list of graphical parameters
    - Set `track.height` as a percentage of plot area.

```{r cd-label2, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05,
             link.sort = TRUE, link.largest.ontop = TRUE,
             annotationTrack = c("grid", "axis"),
             preAllocateTracks = list(track.height = 0.1))

# add labels
circos.track(track.index = 1, bg.border = NA, panel.fun = function(x, y) {
  # create temporary objects for the sector name and x-limits
  reg_lab <- get.cell.meta.data("sector.index")
  xx <- get.cell.meta.data("xlim")
  # use the temporary objects to add text in each sector of the track
  circos.text(x = mean(xx), y = 1, labels = reg_lab, facing = "bending")
})
```



### Labels

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-label2-1.pdf")
```

### Labels

- Still not enough room for longer labels.
    - Increase the `track.height`
    - Create two labels for some regions
    - Reduce the font size using `cex` in `circos.text()` - default is `cex = 1`
```{r}
str_wrap(string = r$reg, width = 14)

r <- r %>%
  # title case for labels
  mutate(lab = str_to_title(string = reg),
         lab = str_replace(string = lab, pattern = "And The", replacement = "&"),
  # use str_wrap to split longer labels into two
         lab = str_wrap(string = lab, width = 14)) %>%
  # separate based on \n
  separate(col = lab, into = c("lab1", "lab2"), sep = "\n", fill = "right") %>%
  # positioning for first lab1, needs to be further out if lab2 exists
  mutate(y = ifelse(test = !is.na(lab2), yes = 1, no = 0.8))
```

### Labels

- Still not enough room for longer labels.
    - Increase the `track.height`
    - Create two labels for some regions
    - Reduce the font size using `cex` in `circos.text()` - default is `cex = 1`
```{r}
r
```


### Labels

```{r cd-label4, message=FALSE, fig.width = 4, fig.height = 4, message=FALSE,  warning= FALSE,  fig.show='hide', fig.path="./plot/"}
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05,
             link.sort = TRUE, link.largest.ontop = TRUE,
             annotationTrack = c("grid", "axis"),
             # increase to 0.2 to fit two lines of labels
             preAllocateTracks = list(track.height = 0.2))

circos.track(track.index = 1, bg.border = NA, panel.fun = function(x, y) {
  s <- get.cell.meta.data("sector.index")
  # filter to row of r for the sector's region to create a temporary rr
  rr <- filter(r, reg == s)
  xx <- get.cell.meta.data("xlim")
  # use temporary rr to add text
  circos.text(x = mean(xx), y = rr$y, labels = rr$lab1, facing = "bending",
              cex = 0.8)
  circos.text(x = mean(xx), y = 0.6,  labels = rr$lab2, facing = "bending",
              cex = 0.8)
})
```

### Labels

\vspace{-0.25cm}
\centering
```{r, echo = FALSE}
knitr::include_graphics("./plot/cd-label4-1.pdf")

```



### Saving

- Always save as PDF to give scalable image
    - We can zoom in very closely and we will still see the chords
    - If we save a vector graphic, e.g. PNG these details will disappear.
- Use the `pdf()` function before the plot to open a PDF
- Use `dev.off()` after the plot code to close the PDF

```{r, message=FALSE, message=FALSE,  warning= FALSE, eval=FALSE}
pdf(file = "./plot/un_stock_2019.pdf", width = 4, height = 4)

circos.par(track.margin = c(0.01, -0.01), gap.degree = 5, start.degree = 90)
chordDiagram(x = d, order = r$reg, grid.col = r$col2, transparency = 0.25,
             directional = 1, direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", diffHeight = -0.05,
             link.sort = TRUE, link.largest.ontop = TRUE,
             annotationTrack = c("grid", "axis"),
             preAllocateTracks = list(track.height = 0.2))
circos.track(track.index = 1, bg.border = NA, panel.fun = function(x, y) {
  s <- get.cell.meta.data("sector.index")
  rr <- filter(r, reg == s)
  xx <- get.cell.meta.data("xlim")
  circos.text(x = mean(xx), y = rr$y, labels = rr$lab1, facing = "bending", cex = 0.8)
  circos.text(x = mean(xx), y = 0.6,  labels = rr$lab2, facing = "bending", cex = 0.8)
})

dev.off()
```

### Saving

- Left:  PNG with `width = 4, height = 4`
- Right: PDF with `width = 4, height = 4`

\begin{center}
\includegraphics[width=0.48\textwidth]{./plot/un_stock_2019.png}
\vspace{0.1cm}
\includegraphics[width=0.48\textwidth]{./plot/un_stock_2019.pdf}
\end{center}

- Could increase resolution of PNG with larger dimensions but at the cost of very large file sizes


### Exercise (ex8.R)

```{r eval = FALSE, prompt=FALSE,  code = readLines('../exercise/ex8.R') }
```

### References




